---
title: "PML Course Project"
author: "The Hawkman"
date: "1/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This document is the final project for Peer Review for Coursera's Practical Machine Learning. It uses two data sets that are partitions of a self-movement data to be analyzed and used to create a training and prediction models.

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#### Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 


#### Load the libraries

```{r}
library(caret)
library(rpart)
library(rpart.plot)
#library(randomForest)

set.seed(5263)

verbose <- TRUE 
```

#### Load the files

```{r}
tfile <- "pml-training.csv"
vfile <-"pml-testing.csv"

########## Check for the file
if(verbose) print("Checking for training file....")
if(!file.exists(tfile))
{
  url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  download.file(url,tfile)
  
}

if(!file.exists(vfile))
{
  url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
  download.file(url,vfile)
  
}

tdf <- read.csv(tfile, sep=",", header=T)
vdf <- read.csv(vfile, sep=",", header=T)

```

#### Set factors, indexes and clean up the data

```{r}
tdf$classe <- as.factor(tdf$classe)


idx <- apply(tdf,2,function(x) {sum(is.na(x))})
tdf <- tdf[,which(idx == 0)]
idx <- apply(vdf,2,function(x) {sum(is.na(x))})
vdf<- vdf[,which(idx == 0)]

v <- which(lapply(tdf, class) %in% "numeric")

```


#### Create the data partition

```{r}

tdp = createDataPartition(tdf$classe, p = 2/3, list=FALSE)
trainDP = tdf[tdp,]
CVDP = tdf[-tdp,]

###take the near zeros out --- will fix "factor kurtosis_roll_belt has new levels" error
###plus takes somewhat irrelevant data out

nzv <- nearZeroVar(trainDP)
trainDP <- trainDP[, -nzv]
CVDP <- CVDP[, -nzv]

mostlyNA <- sapply(trainDP, function(x) mean(is.na(x))) > 0.95
trainDP <- trainDP[, mostlyNA==F]
CVDP <- CVDP[, mostlyNA==F]

trainDP <- trainDP[, -(1:5)]
CVDP <- CVDP[, -(1:5)]
```

#### Create the model and train the data set
I used the ranger package instead of randomForest and it ran much faster.


```{r}
tgrid <- expand.grid(
  .mtry = 2:4,
  .splitrule = "gini",
  .min.node.size = c(10, 20)
)

modFit <- train(classe ~., method="ranger", data=trainDP, trControl=trainControl(method='cv',verboseIter = T, classProbs = T, number=3, allowParallel=TRUE ),
                tuneGrid = tgrid,
                num.trees = 100,
                importance = "permutation")

predictDF <- predict(modFit, trainDP)
confusionMatrix(predictDF, as.factor(trainDP$classe))

cvPred <- predict(modFit, CVDP)
confusionMatrix(cvPred, CVDP$classe)
```

Accuracy is .9963 for the CV Predictor



#### Here we run the model against the test set and display the test set
Make sure to remove the zeros

```{r}
nzv <- nearZeroVar(vdf)
vdf <- vdf[, -nzv]

noZeros <- sapply(vdf, function(x) mean(is.na(x))) > 0.95
trainDP <- vdf[, noZeros==F]
vdf <- vdf

vdf <- vdf[, -(1:5)]

testDF <- predict(modFit, vdf)
```

Display the results

```{r}

testDF
```
```